---
title: "AIC from quasi-possion model"
author: "Meilin Yan"
date: "July 18, 2016"
output: pdf_document
---

# AIC formula
```{r}
# AIC = -2*loglikehood + k*(Number of parameters in the fitted model)

# Number of parameters is the equivalent degree of freedom.
```
  
# Function to compute the Q-AIC in quasi-possion models from Antonio's paper
```{r}
fqaic <- function(model) {
  loglik <- sum(dpois(model$y, model$fitted.values, log=TRUE))
  phi <- summary(model)$dispersion
  qaic <- -2*loglik + 2*summary(model)$df[3]*phi
  return(qaic)
}
```

# Verification of the Q-AIC function
## Fit the model twice with possion (regular likelihood model) and quasi-possion Possion model. Use AIC to choose df per year.

```{r}
library(dlnm)
library(splines)
data("chicagoNMMAPS")
data <- chicagoNMMAPS

cb.pm <- crossbasis(data$pm10, lag=15, argvar=list(fun="lin"),
                    arglag=list(fun="poly", degree=4))

cb.temp <- crossbasis(data$temp, lag=3, argvar=list(df=5),
                      arglag=list(fun="strata", breaks=1))

# Possion model
mod.posi.6 <- glm(death ~ cb.pm + cb.temp + ns(time, 6*14) + dow,
              family = poisson(), data)
mod.posi.7 <- glm(death ~ cb.pm + cb.temp + ns(time, 7*14) + dow,
              family = poisson(), data)
mod.posi.8 <- glm(death ~ cb.pm + cb.temp + ns(time, 8*14) + dow,
              family = poisson(), data)

# Quasi-Possion model
mod.quasi.6 <- glm(death ~ cb.pm + cb.temp + ns(time, 6*14) + dow,
              family = quasipoisson(), data)
mod.quasi.7 <- glm(death ~ cb.pm + cb.temp + ns(time, 7*14) + dow,
              family = quasipoisson(), data)
mod.quasi.8 <- glm(death ~ cb.pm + cb.temp + ns(time, 8*14) + dow,
              family = quasipoisson(), data)
```

## Get log-likelihoods from possion and quasi-possion models
```{r}
# Possion model
loglik.posi <- logLik(mod.posi.6)

# Quasi-Possion model
loglik.quasi <- sum(dpois(mod.quasi.6$y, mod.quasi.6$fitted.values, log=TRUE))

loglik.posi
loglik.quasi
# They are the same
```

## Get number of parameters from the two models
The dispersion parameter, which was forced to be 1 in a possion model, is allowed to be estimated in a quasi-possion model. And the dispersion parameter tells us how many times larger the variance is than the mean. 

+ In a possion model,
number of parameters = number of coefficients.
+ In a quasi-possion model,
number of parameters = (number of coefficients)*(dispersion parameter).

```{r}
# Possion model
# number of coefficients
summary(mod.posi.6)$df[3]
# dispersion parameters
summary(mod.posi.6)$dispersion

# Quasi-Possion model
# number of coefficients
n.coef.quasi <- summary(mod.quasi.6)$df[3]
# dispersion parameters
n.disper.quasi <- summary(mod.quasi.6)$dispersion
# number of parameters
n.coef.quasi*n.disper.quasi
```

## Compare model based on AIC
```{r}
# Possion model
# AIC calculated in R
AIC(mod.posi.6, mod.posi.7, mod.posi.8) # model with 7 df per year has the smallest AIC


# AIC calculated with fqaic function
c(fqaic(mod.posi.6), fqaic(mod.posi.7), fqaic(mod.posi.8)) # model with 7 df per year has the smallest AIC

# Same results but the values of AIC vary a little. I guess R calculates AIC by using a more accurate or complex formula. I am going to figure it out.

# Quasi-possion model
AIC(mod.quasi.6, mod.quasi.7, mod.quasi.8)  # Not avaiable from R

c(fqaic(mod.quasi.6), fqaic(mod.quasi.7), fqaic(mod.quasi.8))   # model with 8 df has the smallest AIC
```

So I think the Q-AIC function is OK for us to extract AIC from a quasi-possion model.


